\textbf{NBA Party}

### 2017-04-09 init_model.rnw
Using continuous measure of party-city. Tidied up formatting. Sent a draft to Guido.

\subsection*{Data}

\paragraph{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house. To this data, we've added features such as 
the number of days since last game, the last game location, the travel distance, and
average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}.

\paragraph{BLS data} We also obtained data from
\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of
establishments by business type at the county level. As a proxy for how much night-life
there is in a city, we look toward the number of sound recording studios, musical groups,
and music publishers there are. For each year and each county, we simply calculate
the average number of establishments across all three music categories listed above.

We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data
for this team.

<<>>=
musicians[, list(team, season, nmusicians)]
@

We then merge this in with our lines data, taking care to do so according to the
last game's location. We also lag our BLS data by one year.

<<>>=
### Merge lines data with BLS data.
lines <- merge(lines, musicians,
               by.x = c('season', 'last.game.loc'),
               by.y = c('season', 'team'), all.x = T)

@@

\paragraph{What cities have nightlife?}

@@

<<>>= 
musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)), by = team]
@

@@

\subsection*{A significant first model} We keep things simple, and try to explain the likelihood
of meeting the spread as a function of whether a team visited a city with an active night-life
the day before.

<<>>=

m <- glm(outcome == 'W' ~ I(nmusicians*(ndays.lgame==1)) + ndays.lgame + travel.dist + avg.age, 
         data = lines[season < 2017 & last.game.loc != team])

@@

The magnitude of our ``\texttt{party}'' variable is both practically and 
statistically significant.

\subsection*{A model worth betting on}
If we re-train an even simpler model, regressing on only the variable of interest, 
we get great betting performance.

@@

>>>>

\section*{Data Sources}

\paragraph{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
<<echo=F>>=
lines[, list(date, team, opponent, line, score)]
@

\paragraph{BLS data} We also obtained data from
\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. \ul{As a 
proxy for how much night-life there is in a city, we look toward the 
number of sound recording studios, musical groups, and music publishers there 
are for a particular county-year}: we simply calculate the average number of establishments across all three music categories listed above.\footnote{We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.}

<<echo=F>>=
musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
musicians[, list(county, team, season, nmusicians = round(nmusicians, 1))]
@

We then merge this in with our lines data, taking care to do so according to the
last game's location. We also lag our BLS data by one year.
To this data, we've created features such as 
the number of days since last game, the last game location, the travel distance, and
average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}.

<<echo=2>>=
### Merge lines data with BLS data.
lines <- merge(lines, musicians,
               by.x = c('season', 'last.game.loc'),
               by.y = c('season', 'team'), all.x = T)

@@

\subsection*{What cities have nightlife?}

@@

<<>>= 
musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)), by = team][order(average.nmusicians)]
@

@@

\subsection*{A significant model} We keep things simple, and try to explain the likelihood
of meeting the spread as a function of whether a team visited a city with an active night-life
the day before.

<<echo=-4>>= 
lines[, party := ifelse(ndays.lgame == 1, nmusicians, 0)]

m <- glm(outcome == 'W' ~ party + ndays.lgame + travel.dist + avg.age, 
         data = lines[season < 2017 & last.game.loc != team])

@@

The sign of our ``\texttt{party}'' variable is intuitively correct, and 
statistically significant.

\subsection*{A model worth betting on}
If we re-train an even simpler model, we obtain fantastic betting performance.

@@

summary(m)$coef %>% round(., 5)
